---
title: "Laboratory work #1"
output:
  word_document: default
  pdf_document: default
always_allow_html: true
---

```{r}
wd <- '~/Documents/r_spatial_data'
setwd(wd)
library(webshot)
library(tinytex)
library(raster)
library(sf)
library('mapview')
```

**Task 1**

To begin with, let's upload the raster 

```{r}
China_population <- raster(paste0('TN1_Donnees/Chine_population.tif'))
China_population
```

a)  Now let's find out raster's dimensions:

```{r}
dim(China_population)
```

... and the number of pixels:

```{r}
ncell(China_population)
```

b) We need to calculate main statistical parameters: maximum value

```{r}
maxValue(China_population)
```

minimum value 

```{r}
minValue(China_population)
```

The mean

```{r}
cellStats(China_population, mean)
```

And the standard deviation 
```{r}
cellStats(China_population, sd)
```

c)  Raster's Coordinate Reference System can be found out as following:

```{r}
crs(China_population)
```

In Proj.4 representation: +proj=moll +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs

d)  The projection used is Mollweid's projection. It belings to the group of pseudocylindrical projections, and it maintains areas

e)  Raster's resolution is:

```{r}
res(China_population)
```

Units are meters (from proj.4 representation: "... +units=m ..." )

f)  Let's calculate some statistics on population density. To be able to do this we need to divide each pixel value by  (2000/1000)\^2 = 4 km\^2 - the area of one pixel

```{r}
China_pop_dens <- China_population/((res(China_population)[1]/1000)^2)
```

Mean population density:

```{r}
cellStats(China_pop_dens, mean)
```

Standard deviation of population density:

```{r}
cellStats(China_pop_dens, sd)
```

g)  China's population is:

```{r}
cellStats(China_population, sum)
```

or 1.37 billion, wich is almost like Wotld Bank's estimation (1.38 billion)

h)  Let's change the resolution to 10\*10 km\^2

```{r}
China_pop_10km <- aggregate(China_population, fact = 5, fun = 'sum')
```

China's population value should be preserved:

```{r}
cellStats(China_pop_10km, sum)
```

Everything line's up

```{r}
China_pop_10km
```

Now we may plot the aggregated raster:

```{r}
plot(China_pop_10km, main = "Spatial distribution of China's population\n2015", axes = F, col = hcl.colors(7, palette = 'Plasma'))
```

**Task 2**

Loading the spatial vector data
```{r}
China_regions_admin <-st_read('TN1_Donnees/Chine_regions_admin.shp')
```

a) Let's find out hte geometry of this data:

```{r}
st_geometry(China_regions_admin)
```

b)  The number of elements

```{r}
length(China_regions_admin$geometry)
```

Attributes
```{r}
length(China_regions_admin)
```

and attributes' names

```{r}
names(China_regions_admin)
```

c) In this layer we can find the following categories of administrative units

```{r}
unique(China_regions_admin$Categories)
```

d) This layer's CRS is:

```{r}
st_crs(China_regions_admin)
```

It's geographical coordinate system. Now let's project our data to Mollweid representation

```{r}
China_regions_admin_moll <- st_transform(China_regions_admin, crs(China_population))
st_crs(China_regions_admin_moll)
```

e) Using mapview() we can combine our raster and vector data and visualize it as an interactive map

```{r}
mapview(China_regions_admin_moll, zcol = 'Categories', lwd = 1, layer.name = 'China administrative regions categories', color = hcl.colors(3, palette = 'Zissou1'), legend.position = c("top", "left")) + mapview(China_pop_10km, layer.name = 'China population', col.regions = hcl.colors(n = 7, palette = 'Oslo'), at = 10^seq(0, 7), legend.position = c("bottom", "right"))

```
